---
title: "TidyTutorial"
author: "Sharon D. Kenny"
date: "October 9, 2019"
output: html_document
r_version: xxxx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the packages I need:
```{r loadLibraries, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

Some functions override others, so to access `filter` from the `stats` package, using this:
`stats::filter()`


## Downloading Dataset

Dataset is here: https://knb.ecoinformatics.org/#view/df35b.304.2  


#### Reading the data

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
head(catch_original)
```


## Selecting our Data

####The pipe operator (%>%)
This operator combines operations, and it is used as a "then"
Keyboard shortcut:  CTRL + SHIFT + m


Step 1.  Remove the "All" column by selecting the columns you want with `select()`.  You can list the columns you would like to have in your new database or... 

```{r}
catch_data <- catch_original %>%  
              select(-All, - notesRegCode)
head(catch_data)
```

Step 2.  Check if the data is right

```{r}
summary(catch_data)
```

Step 3. Chinook is suposed to be a numeric, so let's change the values from character using `mutate()`

```{r}
catch_clean <- catch_data %>% 
                  mutate(Chinook = as.numeric(Chinook))
head(catch_clean)
```

Step 4. Investigate this NAs by using `which()` and `is.na()`


```{r}

i <- which(is.na(catch_clean$Chinook))
i
# This is the index of the row that has the NA
```

We should look into the original copy to vieww that 401th row

```{r}
catch_original[i,]

```

Cool... it is an uppercase i.  This should be a one! Cause: OCR with PDF 

Step 5. Mutate the I into 1

```{r}
catch_clean <- catch_data %>% 
                  mutate(Chinook = ifelse((Chinook == "I"), 1, Chinook)) %>% 
                  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```


You can also write `ifelse (Chinook %in%  c("I" , 'l'), 1, Chinook)`


## Tidy data

#### Pivot your data.  Move from the wide format to tall format using `pivot_longer`.  It is similar to `gather`, which is deprecated.

```{r}
catch_long <- catch_clean %>%
                pivot_longer(cols = -c(Region, Year), 
                             names_to = 'species',
                             values_to = 'catch')
head(catch_long)
  
```


But the wide format is better for printing...

```{r}
catch_wide <- catch_long %>% 
                pivot_wider(names_from = species,
                            values_from = catch)

head(catch_wide)
```

### The way done at the class  
Let's rename our column catch to `catch_thousands` and use `mutate()` 
Well.. not rename it because `catch` is already in thousands.

```{r}
# catch_long <- catch_long %>%  
#                 rename(catch_thousands = catch) %>% 
#                 mutate(catch_thousands = catch_thousands *1000)

catch_long <- catch_long %>%  
                mutate(catch = catch *1000)

head(catch_long)
```


### The way I would like to do it instead:

```{r}
catch_long2 <- catch_long %>%  
                mutate(catch_thousands = catch *1000)

head(catch_long2)
```



## Analyzing Data = Summarize


Mean catch by region


```{r}
mean_region <- catch_long %>% 
                  group_by(Region) %>% 
                  summarise(mean_catch = mean(catch))

mean_region
```



Let's filter by Region & species first if we want to look only at Sockeye

```{r}
mean_region <- catch_long %>% 
                  group_by(Region, species) %>% 
                  summarise(mean_catch = mean(catch),
                            n_obs = n())

mean_region
```

Mean sockeye by region... let's filter first

```{r}
mean_Sockeye_region <- catch_long %>% 
                  filter (species == 'Sockeye') %>% 
                  group_by(Region ) %>% 
                  summarise(mean_catch = mean(catch))

mean_Sockeye_region
```

Now by year





```{r}
annual_catch <- catch_long %>% 
                  group_by(Year) %>% 
                  summarise(total_catch = sum(catch)) %>% 
                  arrange(desc(total_catch))

annual_catch
```

## Joins in DPLYR
#### Joining by region column


```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                            stringsAsFactors = FALSE) %>% 
                            select(code, mgmtArea)

head(region_defs)
```

Important: Create a new table when joining instead of overwriting a dataframe with the joined rows

Below we will be able to "see" the left table and the right table

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

or

```{r}
region_defs2 <- region_defs %>% 
                    rename(Region = code)

catch_joined <- left_join(catch_long, region_defs2, by = "Region")
# This is another way, with an inner join, but it will be different
# catched_joined <- left_join(catch_long, region_defs)

region_defs2
```


#### Separate and Unite sample

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df 


dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE) %>% 
  unite(date_2, year, month, day, sep = "-")
```


Lubridate is a great way to work with dates!
