---
title: "R Intro: DataScience @ SantaBarbara"
author: "Sharon D. Kenny"
date: "October 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### R Markdown Tutorial

1. Details on using R Markdown see <http://rmarkdown.rstudio.com>.
      a. When you click the **Knit** button
      b. Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.  


###### First chunk below  

  

```{r firstChunk}
(x = "Hello world")

# Inside the chunk, the hash is a comment

```




#### Python! Hurray! :)

```{r}
library(reticulate)
```


```{python, eval = TRUE}
x = '3'
y = x*3
print('This is y:' y)

# It works!!!
```


Plain text with two spaces at the end...  
... to start a new paragraph

```{r}
xx <- 2.222222222
```

The result is `r xx`


> This is a block quote.  :)

```{r, warning = FALSE}
t <- c('1', '2', 'gggg')
t <- as.numeric(t)
t
```

Inline equations with single dollar sign
$$ \sum(x^n) $$


x = "create this as a quote by selecting all and typing a quote"



# Read a file / dataframe
Read in a file downloaded from the [Arctic Data Center](https://doi.org/10.18739/A25T3FZ8X).

```{r}
bgCHEM <- read.csv('BGchem2008data.csv', stringsAsFactors = FALSE)
# head(bgCHEM)
```

Let's select some columns with `$`, which is a list selector 

```{r}
meanTemp <- mean(bgCHEM$CTD_Temperature)
# meanTemp
```



```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's use the base-R plot 

```{r}
plot(bgCHEM$CTD_Depth,bgCHEM$CTD_Temperature)
```

