---
title: "Alaskan Salmon"
author: "Sharon D. Kenny"
date: "October 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, warning=FALSE, message = FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(sf)
library(ggmap) # devtools::install_github("dkahle/ggmap")
library(DT)
library(leaflet)
```


Escapment daily counts (number of salmon fish that jumped upstream)

```{r, results = 'hide'}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = FALSE)

head(esc)
```


### Calculate Annual Escapment for Salmon Species  

echo = false means that the code will not show

```{r, echo=FALSE}
annual_esc <- esc %>% 
  separate (sampleDate, c("Year", "Month", "Day"), sep = '-', remove = FALSE) %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Year, SASAP.Region, Species) %>% 
  summarise(total_escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Pink", "Coho" ))


head(annual_esc)
```


The same as above:

filter (Species == "Chinook" | Species == "Sockeye")

Where the pipe means `or`


```{r}
#After giving `aes`, you have to give the geometry to ggplot

#One way to color your graph:
#ggplot(data = annual_esc, mapping = aes(x = Species, y= total_escapement)) +
#  geom_col(fill = 'blue')

ggplot(data = annual_esc, mapping = aes(x = Species, y= total_escapement, fill = SASAP.Region)) +
  geom_col()

```

#### A different plot...

```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = Year, y = total_escapement, color = Species)) + 
    geom_line() +
    geom_point() +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_bw()
```


#### Creating plots for every region using facet_wrap


```{r}
ggplot(annual_esc, aes(x = Year, y = total_escapement, color = Species)) + 
    geom_line() +
    geom_point() +
    facet_wrap(~SASAP.Region, scales = "free_y") +
    ylab("Escapement") +
    theme_bw()
```

Escapement over time: all regions
```{r}
fig1 <- ggplot(annual_esc, aes(x= Year, y=total_escapement, color = Species)) +
  geom_point() +
  geom_line()  +
  facet_wrap(~SASAP.Region, scales = 'free_y', ncol = 1) +
    theme_bw()

fig1

ggsave("escapement.png", fig1, width =12, height = 7, units = 'in')
```

# Map Sampling Locations

```{r}
locations <- esc %>% 
  distinct (Location, Latitude, Longitude)

head(locations)

```

disctinct will grab uniques values from one or more columns
 
Let's drop the na's

```{r}
locations <- esc %>% 
  distinct (Location, Latitude, Longitude) %>% 
  drop_na()

head(locations)
```




```{r}
datatable(locations)

```


View is different... it does not knit  

```{r}
View (locations)
```

Let's map.  Hurray!

The tilde works as the `aes` function

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers( lng = ~ Longitude, lat = ~ Latitude, popup = ~ Location)

```


### Working with tiles


esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = FALSE)

```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
        layers = 'GEBCO_LATEST',
        attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~ Longitude,
                   lat = ~ Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```



SDK




